---
title: "EDA"
output:
  html_document:
    toc: yes
    code_folding: hide
    df_print: paged
  pdf_document:
    toc: yes
---

```{r}
knitr::opts_chunk$set(message=FALSE, warning = FALSE, tidy.opts=list(width.cutoff=55))
```

```{r}
library(tidyverse)
library(dplyr)
library(mosaic)  # standardizing variables
library(ggplot2)
library(ggcorrplot)
library(GGally)
```

# Load the Data

```{r}
# Set the correct path to the file
file_path <- "./data/healthcare-dataset-stroke-data.csv"

# Load the dataset
data <- read.csv(file_path)
```

# Inspect the Data

## Variable Information:

```{r}
# View the first few rows of the dataset
head(data)

# structure
str(data)

# Summarize the dataset to get a quick overview
summary(data)

```


* id: unique identifier
* gender: 'Male', 'Female' or 'Other'
* age: age of the patient
* hypertension: 0 if the patient doesn't have hypertension, 1 if the patient has hypertension
* heart_disease: 0 if the patient doesn't have any heart diseases, 1 if the patient has a heart disease
* ever_married: 'No' or 'Yes'
* work_type: 'children', 'Govt_jov', 'Never_worked', 'Private' or 'Self-employed'
* Residence_type: 'Rural' or 'Urban'
* avg_glucose_level: average glucose level in blood
* bmi: body mass index
* smoking_status: 'formerly smoked', 'never smoked', 'smokes' or 'Unknown'
* stroke: 1 if the patient had a stroke or 0 if not


```{r}

# Check for missing values
colSums(is.na(data))
```
# EAD

Visualizes the dataset's age distribution using a histogram, and generates a correlation matrix plot for numeric columns.
```{r}
#drop ID
data_ana <- data %>% select(-id)
data_ana$bmi <- as.numeric(factor(data$bmi))-1

# Plotting age distribution
ggplot(data_ana, aes(x=age)) + geom_histogram(bins=30, fill='blue', color='black') + ggtitle('Age Distribution')

# Plotting correlation matrix
corr <- cor(select_if(data_ana, is.numeric))
corr
#plot the correlation matrix
ggcorrplot(corr, method='circle', hc.order=FALSE, type='lower', lab=TRUE)

```

```{r}
# Boxplot for comparing bmi across different stroke outcomes
ggplot(data, aes(x = as.factor(stroke), y = bmi, fill = as.factor(stroke))) +
  geom_boxplot() +
  labs(x = "Stroke (1 = Yes, 0 = No)", y = "BMI") +
  ggtitle("BMI Distribution by Stroke Outcome")
```